Secure Multi-Party Computation with SCALE-MAMBA

### 2019-04-19 paper.tex
Freitag abend

\begin{abstract}
This paper takes a look at Secure Multi-Party Computation(SMPC), especially using SCALE-MAMBA. For that, it introduces first what SMPC is, how SPDZ, the predecessor of SCALE-MAMBA, works and what are the differences. It discusses the functionalities of SCALE-MAMBA, especially the split into online and offline phase, how the beaver tripels are used for sacrificing and in which complexity classes the calculations are done. Finally it compares scale mamba shortly to other competitors in the market and gives an outlook what next technologies are coming right now and are coming in the near future in SMPC.
\end{abstract}

\section{Motivation and Introduction}\label{sec:intromoti}

In todays world, and even more in the future, there are a lot of computations that have to be done in a distributed environment, because in a lot of use cases multiple different parties are involved, but also because the amount of data is getting so big that it often becomes impossible to have a single entity that processes and stores all the data. Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive information or because they do not want to share information with a lot of business value.\\
A basic example is an anonymous voting. In most elections there is a third party that has to aggregate and count the votes made by the voters. For instance, in Germany's elections this is done by specific people counting a small amount of votes and then forwarding the information to a central institution. The other problem is that it has to be guaranteed that a vote is valid, meaning that each voter holds the right to vote and is only voting one time. So if there would be a malicious person in this specific system, it would be easy for that person to gather information over a small subset of voters or even manipulate the local results.\\
Another use case is the exchange of medical data. Medical data of patients is highly confidential and should never be published. But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it is helpful to compare the symptoms to the diagnoses of other doctors.\\
Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
The basic problem is how to evaluate a common function, with multiple parties having different inputs to this function, without revealing the different inputs of the parties. Although there are a lot of theoretical solutions for this problem, real-world implementations are still not performing on the desired level. So in the last years the research shifted more into the area of how to make the existent solutions performant enough while still guaranteeing a desired level of security.\\
This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the Katholieke Universiteit Leuven, which is one of the forerunners in SMPC and a system that combines a lot of research by different teams with the newest ongoing research to provide a platform with highly performant SMPC. Nevertheless, this paper will at first give a short introduction into the history of SMPC and will try to explain its basic technologies to the reader. It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparison to other SMPC systems which use different approaches, especially considering the usability in different use cases. Ultimately, it will try to give an outlook what might be coming in the next years.\\

@@

GENERATION OF BEAVER TRIPLES

@@

\subsubsection{Upcoming changes to SCALE-MAMBA}
SCALE-MAMBA is a research system, so there will be a lot more research time spend on improving the system further. An important aspect is the usability in the practice. Because online and offline phase are integrated into each other, the complete runtime sometimes takes a lot of time. This is especially the case in full threshold access structures. Because of the integration, it is right now not possible to execute the offline phase earlier. One of the currently ongoing improvements on SCALE-MAMBA tackles exactly this problem. The developer team tries to give MAMBA the possibility to compile code just in time. That would offere the possibility, that the offline data is calculated before and then, when the need arises, a computatation could be specified in MAMBA and then intime compiled and executed, without having to wait for the expensive offline calculations.\\

\section{Implementations}
Because SCALE-MAMBA was only released in 2018, there are not a lot documented usecases in praxis. Nevertheless, because of the structural and alogirthmic similarities to the earlier versions of SPDZ, taking a look on SPDZ implementations should give a general understanding of the possible usage of SCALE-MAMBA.\\

\subsection{SPDZ 2.0 in medical healthcare}
In medical healthcare, preserving privacy plays a huge role because patient records hold highly sensitive data, thus, its a natural area for the application of privacy protecting technologies. In [TODO: REFERENCE] a clinical decision support system was implemented with SPDZ 2.0. The idea is that the system provides the optimal treatment for a patient with HIV based on the  records of similar patients in different hospitals. To preserve privacy, the hospitals compute the efficiency of one treatment for one patient together with SMPC.\\
The researchers used a setup of 20.000 patient records split between the hospitals, and were able to compute the effectiveness of 100 treatments for one patient in 24 minutes in the online phase. For calculating the effectiveness of one treatment, their algorithm needed 40 million multiplications. They only implemented the online phase of SPDZ, but showed that it scaled linearly with the amount of patient records.

TODO: CHECK HOW MANY HOSPITALS

>>>>

\begin{abstract}
This paper examines Secure Multi-Party Computation(SMPC), especially the SCALE-MAMBA protocol. To do that, it first introduces SMPC, afterwards SPDZ, the predecessor of SCALE-MAMBA, and, after taking a detailed look at SCALE-MAMBA, compares them. It focuses on the algorithmic foundations of SCALE-MAMBA, especially the split into online and offline phase, how beaver triples are used to enable computational efficient multiplications and which computational complexities the algorithms posses. Afterwards, it tries to depict the evolution in the different versions of the protocol and to locate SCALE-MAMBA between its competitors. Ultimately, it will show some exemplary implementations in real world applications and will give an outlook what further evolutions might be coming in the next years.
\end{abstract}

\section{Motivation and Introduction}\label{sec:intromoti}
In today's world, and even more in the future, there exist a lot of computations that have to be done in a distributed environment, because in a lot of use cases multiple parties are involved, but also because the amount of data is getting so big that it often becomes impossible to have a single entity that processes and stores all of it. Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive or it has a high business value.\\
A basic example is an anonymous voting. In most elections there is a third party that has to aggregate and count the votes. For instance, in Germany's elections[TODO: LINK] this is done by specific people counting a small amount of votes and then forwarding the information to a central institution. The other problem is that it has to be guaranteed that a vote is valid, meaning that each voter holds the right to vote and is only voting one time. So in an ideal system, no person participating in the aggregation process should be able to get any information over any vote or any subset of voters.\\
Another use case is the exchange of medical data. Medical data of patients is highly confidential and should never be published. But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it would be helpful to compare the symptoms to the diagnoses of other doctors.\\
Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
The basic problem is the evaluation of a common function, to which multiple parties have different inputs, without revealing these inputs. Although there are a lot of theoretical solutions for this problem, real-world implementations are still not performing on the desired level. So in the last years the research shifted more into the area of how to improve the performance of existing solutions while still guaranteeing a desired level of security.\\
This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the Katholieke Universiteit Leuven. This system is one of the forerunners in SMPC and combines a lot of research by different teams to provide a platform with highly performant SMPC. Nevertheless, this paper will at first give a short introduction into the history of SMPC and will try to explain its basic technologies to the reader. It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparison to other SMPC systems which use different approaches, especially taking a look on which different requirements are fulfilled by each. Ultimately, it will try to give an outlook what might be coming in the next years.\\

@@

GENERATION OF BEAVER TRIPLES: Include protocol on how triples are distributively generated without revealing them

@@

\subsubsection{Upcoming changes to SCALE-MAMBA}
SCALE-MAMBA is a research system, so there will be a lot more research time spend on improving the system further. An important aspect is the usability in the practice. Because online and offline phase are integrated into each other, the complete runtime sometimes takes a lot of time. This is especially the case in full threshold access structures. Because of the integration, it is right now not possible to execute the offline phase earlier. One of the currently ongoing improvements on SCALE-MAMBA tackles exactly this problem. The developer team tries to give MAMBA the possibility to compile code just in time as mentioned in [TODO: link to documentation]. That would offer the possibility, that the offline data is calculated before and then, when the need arises, a computation could be specified in MAMBA and then in-time compiled and executed, without having to wait for the expensive offline calculations.\\

\section{Implementations}
Because SCALE-MAMBA was only released in 2018, there are not a lot documented use cases in praxis. Nevertheless, because of the structural and algorithmic similarities to the earlier versions of SPDZ, taking a look on SPDZ implementations should give a general understanding of the possible usage of SCALE-MAMBA.\\

\subsection{SPDZ 2.0 in medical healthcare}
In medical healthcare, preserving privacy plays a huge role because patient records hold highly sensitive data, thus, its a natural area for the application of privacy protecting technologies. In [TODO: REFERENCE] a clinical decision support system was implemented with SPDZ 2.0. The idea is that the system provides the optimal treatment for a patient with HIV based on the  records of similar patients in different hospitals. To preserve privacy, the hospitals compute the efficiency of one treatment for one patient together with SMPC.\\
The researchers used a setup of 20.000 patient records split between the hospitals, and were able to compute the effectiveness of 100 treatments for one patient in 24 minutes in the online phase. For calculating the effectiveness of one treatment, their algorithm needed 40 million multiplications. They only implemented the online phase of SPDZ, but showed that it scaled linearly with the amount of patient records. In the paper the amount of parties participating in the computation was not specified, but because the computation is quite good, it can be assumed that a really simple setup is used that doesn't create a lot of overhead.

TODO: compare, what is the speed and how fast are different ones? 
TODO: CHECK HOW MANY HOSPITALS
